{\bf {\Large Evaluation and correction of noise related to physical processes in estimates of estuary metabolism}}

### 2014-08-14 swmp_noise.tex
finished tables/figs for simulation assessment after detiding worked

Multiple time series were simulated by varying the conditions in each of the above equations (\cref{fig:sim_ex}). Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Additionally, five window widths for decimal time in the weighted regression were evaluated: 2, 10, 20, 30, and 40 days.  In total, five unique window widths were evaluated for each of 243 simulated time series, producing results for 1215 weighted regressions.

The detided values for each regression were compared to the biological \ac{DO} signals in the simulated data to evaluate the method. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions sufficiently detided the \ac{DO} time series for all simulations and window width combinations (\cref{tab:dtd_perf}). Mean correlation for all time series and window widths between the detided and biological values was 0.99, with values ranging from 0.70 to 1.00.  Mean error was 0.16, with values ranging from 0.00 to 1.04.  Minimal variation in model performance was observed with different characteristics of the \ac{DO} time series and window widths, such that results were generally satisfactory (i.e., high correlations, low errors) for all simulations (\cref{tab:dtd_perf,fig:err_surf}).  Ability of the weighted regressions to detide the \ac{DO} signal increased with decreasing amplitude of the diel \ac{DO} component and increasing window widths.  Results were not, or only minimally, affected by changes in the effects of tidal advection ($DO_{adv}$) and variation in the magnitudes of process and observation uncertainty.  The models were also minimally affected by changes in tidal category, with slightly better performance for semidiurnal tides.

A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.  In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).  Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.  An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.  \Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).  Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.  These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.  As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.  Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.  Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.

@@

@@

@@

\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.  Simulated values were estimated every thirty minutes for 31 days.  Yellow indicates daylight periods.\label{fig:do_sim}}
\end{figure}

@@

% example of detiding a simulated time series
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: could not find function "{}wtreg\_fun"{}}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: object 'prd\_tmp' not found}}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error: it is only valid to get a child from a "{}gTree"{}}}\end{kframe}
\end{knitrout}
\vfill
\clearpage

@@

{\centering \includegraphics[width=\maxwidth]{figure/err_surf}

@@

\caption[Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results]{Examples of correlations (left) and errors (right) between detided and biological \ac{DO} from weighted regression results.  Each tile represents a correlation or error value from results for a given combination of simulation parameters ($DO_{adv}$, $DO_{die}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and selected window widths (two and ten days).  Results are for simulations using a diurnal tidal time series.\label{fig:err_surf}}
\end{figure}

@@

% example of extreme scenario
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/extreme}

@@

\caption[An extreme example of results from the weighted regression with decreasing window widths]{An extreme example of results from the weighted regression with decreasing window widths.  Results are for a simulated time series with a diurnal tidal component, no effect of tidal advection ($DO_{adv} = 0$), and an observed \ac{DO} signal dominated only by a diel component ($DO_{bio} = 2$, $\epsilon_{obs} = 0$, $\epsilon_{pro} = 0$).  Window widths varied form 2 to 40 days.  See \cref{do_tid,do_mtd,do_dtd,do_obs} for notation.\label{fig:extreme}}
\end{figure}

@@

@@

% summary of simulation performance for detided and biological
% latex.default(tab, file = "", rowlabel = "Parameter", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = c(rep(3,          5), 5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(2,          2), rowname = rows, colheads = rep(c("Mean", "Std. Err."),          2), label = "tab:dtd_perf") 
%
\begin{table}[!tbp]
\caption{Mean correlations and error estimates between detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$) and window widths (days).  Each value represents a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf}} 
\begin{center}
\begin{tabular}{lllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{2}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries RMSE}\tabularnewline
\cline{2-3} \cline{5-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}\tabularnewline
\hline
{\bfseries Tidal category}&&&&&\tabularnewline
~~Diurnal&0.97&0.02&&0.24&0.09\tabularnewline
~~Semidiurnal&1.00&0.00&&0.07&0.02\tabularnewline
~~Mixed Semidiurnal&0.99&0.01&&0.18&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&\tabularnewline
~~0&1.00&0.00&&0.07&0.02\tabularnewline
~~1&0.98&0.01&&0.15&0.05\tabularnewline
~~2&0.97&0.02&&0.27&0.10\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&\tabularnewline
~~0&0.99&0.01&&0.16&0.07\tabularnewline
~~1&0.99&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.16&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&\tabularnewline
~~0&0.98&0.02&&0.16&0.07\tabularnewline
~~1&0.98&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.17&0.07\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&\tabularnewline
~~0&0.98&0.02&&0.13&0.07\tabularnewline
~~1&0.99&0.01&&0.16&0.07\tabularnewline
~~2&0.99&0.01&&0.21&0.07\tabularnewline
\hline
{\bfseries Window}&&&&&\tabularnewline
~~2&0.95&0.02&&0.38&0.10\tabularnewline
~~10&0.98&0.01&&0.24&0.08\tabularnewline
~~20&1.00&0.00&&0.09&0.02\tabularnewline
~~30&1.00&0.00&&0.05&0.01\tabularnewline
~~40&1.00&0.00&&0.04&0.01\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

@@

% correlations with tide before/after wtreg
% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "DO flux", "Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res") 
%
\begin{table}[!tbp]
\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations, instantaneous \ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.  \ac{DO} values are correlated with predicted tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.  Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.  Results are for weighted regressions using a window of thirty days and tidal height proportion of one.\label{tab:cor_res}} 
\begin{center}
\begin{tabular}{llllll}
\hline\hline
\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{DO flux}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&\tabularnewline
~~Observed& 0.47***& 0.26***& 0.60***& 0.73***& 0.35***\tabularnewline
~~Detided& 0.01 &-0.04***&-0.08 &-0.17**&-0.01 \tabularnewline
\hline
{\bfseries PDBBY}&&&&&\tabularnewline
~~Observed&-0.45***&-0.31***&-0.33***&-0.46***&-0.24***\tabularnewline
~~Detided&-0.02*& 0.01 & 0.23***& 0.12*&-0.10 \tabularnewline
\hline
{\bfseries RKBMB}&&&&&\tabularnewline
~~Observed& 0.28***& 0.39***& 0.35***& 0.40***& 0.26***\tabularnewline
~~Detided&-0.01 & 0.02**&-0.23***&-0.20***& 0.01 \tabularnewline
\hline
{\bfseries SAPDC}&&&&&\tabularnewline
~~Observed& 0.48***& 0.62***& 0.53***& 0.70***& 0.40***\tabularnewline
~~Detided&-0.01 &-0.03***& 0.04 & 0.15**&-0.02 \tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

>>>>

Multiple time series were simulated by varying the conditions in \cref{do_obs,do_bio,do_unc,do_obs_all,do_sin,do_unc_n,deltdo,deltx,do_advp,do_adv}. Specifically, the simulated data varied in the relative amount of noise in the measurement, relative amplitude of the diel \ac{DO} component, degree of association of the tide with the \ac{DO} signal, and tidal type as diurnal, semidiurnal, and mixed semidiurnal.  Three levels were evaluated for each variable: relative noise as 0, 1, and 2 standard deviations for both process and observation uncertainty, amplitude of diel biological \ac{DO} as 0, 1, and 2 mg L$^{-1}$, and \ac{DO} change from tidal advection as 0, 1, and 2 mg L$^{-1}$.  Three tidal categories were created from \cref{do_sin} using a period of 24.82 hours (principal lunar) for diurnal, 12.42 hours (principal lunar semidiurnal) for semidiurnal, and adding both diurnal and semidiurnal series for mixed semidiurnal. A total of 243 time series were created based on 81 unique combinations of parameters for each tidal category (\cref{fig:sim_ex}).  Three window widths for each variable used to estimate the total weight vetor were evaluated (as half window widths): decimal time as 1, 3, and 8 days, time of day as 6, 12, and 24 hours, and tidal height as 0.25, 0.5, and 1 as a proportion of the observed value at the center of the window.  In total, 27 window width combinations were evaluated for each of 243 simulated time series, producing results for 6561 weighted regressions.

The detided or normalized values for each regression were compared to the simulated data to evaluate the ability of weighted regression to reproduce the biological \ac{DO} signals. Results were summarized using Pearson correlation coefficients and the \ac{RMSE} between the predicted and observed \ac{DO} values and the detided and biological \ac{DO} values.  Overall, the weighted regressions provided accurate results for the detided time series compared to the `true' biological time series regardless of the simulation parameters (\cref{tab:dtd_perf1}) or window widths (\cref{tab:dtd_perf2}).  Mean correlation for all time series and window widths between the detided and biological values was NA, with values ranging from -0.78 to 1.00.  Mean error was 1.27, with values ranging from 0.00 to 2.47.  Simulations with very poor performance (e.g., negative correlations) were those that had minimum widths for day windows and maximum widths for hour windows, or were those with the \ac{DO} signal composed entirely of noise from observation uncertainty. Characteristics of \ac{DO} time series that contributed to improved model performance were increasing amplitude of the diel \ac{DO} component ($DO_{die}$) and increasing process uncertainty ($e_{pro}$), whereas increasing observation uncertainty contributed to decreasing performance (\cref{tab:dtd_perf1}).  Model performance was not related to magnitude of the tidal advection component ($DO_{adv}$), whereas models with diurnal tidal components had slightly decreased performance.  Increasing widths for day and tidal proportion windows contributed to increasing model performance, whereas the opposite was true for increasing hour windows (\cref{tab:dtd_perf2}).

%%%%%
A closer examination of the results for unique scenarios can provide additional insight into characteristics of observed \ac{DO} that influence model results.  In particular, the performance of the model depends entirely on the ability to predict variation in \ac{DO} from tidal effects ($DO_{tid}$, \cref{fig:do_dtd}) and the mean response of \ac{DO} conditional given constant tidal height ($DO_{mtd}$, \cref{fig:do_dtd}).  Specific characteristics of the \ac{DO} time series (e.g., relative magnitude of $\epsilon_{proc}$, $DO_{adv}$, etc.) influence the ability of the model to predict unique sources of variation in observed \ac{DO}.  An extreme case is presented by time series that have no influence of tidal advection (i.e., $DO_{adv} = 0$) on the observed time series.  \Cref{fig:extreme} shows the model results from a simulated time series with no tidal component and the observed \ac{DO} signal is composed only of a diel component ($DO_{bio} = 2$).  Although the `correct' results were obtained with large window widths, the model predicts increasingly periodic contributions of the tidal ($DO_{tid}$) and mean tidal component ($DO_{mtd}$) with decreasing window widths.  These results can be explained by phase synchrony or asynchrony between the simulated tidal component and observed \ac{DO} as the model fits the time series to a non-existent tidal advection component.  As window width decreases, higher importance is given to observations within the window such that periods with high synchrony or asynchrony between observed \ac{DO} and the tidal component will have a larger influence on the model parameters.  Although application of the model to such a time series would be impractical, the results illustrate the importance of preliminary evaluations of the data and considerations for window widths.  Time series with no obvious effect of tidal advection should use relatively large window widths because results from small window widths may be impractical.

@@

 Assumptions of relationships between variables are not necessary using this adaptive weighting scheme given that the method has the ability to interpret statistical relationships without prior knowledge.

@@

\section{Figures}

% example of detiding a simulated time series
\centering\vspace*{\fill}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[!ht]

{\centering \includegraphics[width=\maxwidth]{figure/do_dtd}

\caption[Example of detiding a simulated \ac{DO} time series]{Example of detiding a simulated \ac{DO} time series.  Simulated values are those in \cref{fig:do_sim}.  Yellow indicates daylight periods.\label{fig:do_dtd}}
\end{figure}

@@

\caption[Example of creating a simulated \ac{DO} time series]{Example of creating a simulated \ac{DO} time series.  Values were simulated every 30 minutes for 30 days.  Yellow indicates daylight periods.\label{fig:do_sim}}
\end{figure}

@@

@@

{\centering \includegraphics[width=\maxwidth]{figure/err_surf1}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying simulation parameters: tidal category (e.g., Mixed), strength of tidal association with \ac{DO} signal $DO_{adv}$, amount of process uncertainty $\epsilon_{pro}$, amount of observation observation uncertainty $\epsilon_{obs}$, and strength of diel \ac{DO} component $DO_{die}$.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of simulation parameters.  Results are averaged for all window widths used to evaluate the regressions.\label{fig:err_surf1}}
\end{figure}

@@

{\centering \includegraphics[width=\maxwidth]{figure/err_surf2}

@@

\caption[Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths]{Correlations and errors (\ac{RMSE}) for detided \ac{DO} time series, $DO_{dtd}$, from weighted regression with `true' biological \ac{DO}, $DO_{bio}$, for varying half window widths: days, hour of day, and proportion of tidal range.  Each tile represents the correlation or error between detided and biological \ac{DO} time series from results for a given combination of window widths.  Results are averaged for all simulation parameters used to evaluate the regressions.\label{fig:err_surf2}}
\end{figure}

@@

% example of extreme scenario
\centering\vspace*{\fill}

@@

\section{Tables}

% summary of simulation performance for detided and biological, sim parameters
% latex.default(tab, file = "", where = "h", rowlabel = "Parameter",      caption = cap.val, caption.loc = "top", rgroup = Parms, n.rgroup = rep(3,          5), cgroup = c("Correlation", "RMSE"), n.cgroup = c(5,          5), rowname = rows, colheads = rep(c("Min", "25\\textsuperscript{th}",          "Mean", "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf1") 
%
\begin{table}[h]
\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for different simulation parameters (tidal category, $DO_{die}$, $DO_{adv}$, $\epsilon_{pro}$, $\epsilon_{obs}$).  Values represent a combination of results from multiple simulations with the parameter value held constant for each row (e.g., row one is a summary of all simulations for which the tidal category was diurnal).\label{tab:dtd_perf1}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Parameter}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries Tidal category}&&&&&&&&&&&\tabularnewline
~~Diurnal&-0.78&0.35&0.54&0.80&1.00&&0.00&0.86&1.29&1.99&2.39\tabularnewline
~~Semidiurnal&-0.28&0.37&0.58&0.84&1.00&&0.00&0.73&1.25&1.97&2.40\tabularnewline
~~Mixed Semidiurnal&-0.39&0.37&0.57&0.83&1.00&&0.00&0.78&1.27&1.97&2.47\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{die}}$}&&&&&&&&&&&\tabularnewline
~~0& 0.00&0.21&0.52&0.92&1.00&&0.00&0.26&1.05&1.96&2.05\tabularnewline
~~1&-0.78&0.36&0.55&0.74&1.00&&0.12&0.68&1.23&2.00&2.13\tabularnewline
~~2&-0.78&0.48&0.62&0.81&1.00&&0.25&1.13&1.53&2.08&2.47\tabularnewline
\hline
{\bfseries $\boldsymbol{DO_{adv}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
~~1&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
~~2&-0.78&0.36&0.56&0.82&1.00&&0.00&0.78&1.27&1.98&2.47\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{pro}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.14&0.50&0.83&1.00&&0.00&0.74&1.23&1.96&2.45\tabularnewline
~~1& 0.14&0.35&0.55&0.81&0.99&&0.07&0.76&1.26&1.98&2.45\tabularnewline
~~2& 0.33&0.47&0.65&0.82&0.99&&0.15&0.82&1.32&2.00&2.47\tabularnewline
\hline
{\bfseries $\boldsymbol{\epsilon_{obs}}$}&&&&&&&&&&&\tabularnewline
~~0&-0.78&0.80&0.84&0.97&1.00&&0.00&0.26&0.57&0.78&1.57\tabularnewline
~~1&-0.07&0.39&0.51&0.67&0.84&&0.96&1.01&1.18&1.25&1.82\tabularnewline
~~2& 0.00&0.22&0.34&0.44&0.63&&1.93&1.98&2.07&2.11&2.47\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

% summary of simulation performance for detided and biological, window widths
% latex.default(tab, file = "", rowlabel = "Window", caption = cap.val,      caption.loc = "top", rgroup = Parms, n.rgroup = rep(3, 3),      cgroup = c("Correlation", "RMSE"), n.cgroup = c(5, 5), rowname = rows,      colheads = rep(c("Min", "25\\textsuperscript{th}", "Mean",          "75\\textsuperscript{th}", "Max"), 2), label = "tab:dtd_perf2") 
%
\begin{table}[!tbp]
\caption{Summary (range, mean, percentiles) of correlations and error estimates comparing detided and biological \ac{DO} time series for simulations using different half window widths in the weighted regressions (days, hours, and proportion of tidal range).  Values represent a combination of results from multiple simulations with the window value held constant for each row (e.g., row one is a summary of all simulations for which the half window width was one day).\label{tab:dtd_perf2}} 
\begin{center}
\begin{tabular}{llllllclllll}
\hline\hline
\multicolumn{1}{l}{\bfseries Window}&\multicolumn{5}{c}{\bfseries Correlation}&\multicolumn{1}{c}{\bfseries }&\multicolumn{5}{c}{\bfseries RMSE}\tabularnewline
\cline{2-6} \cline{8-12}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Min}&\multicolumn{1}{c}{25\textsuperscript{th}}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{75\textsuperscript{th}}&\multicolumn{1}{c}{Max}\tabularnewline
\hline
{\bfseries Days}&&&&&&&&&&&\tabularnewline
~~1&-0.78&0.30&0.53&0.77&1.00&&0.00&0.95&1.29&1.96&2.40\tabularnewline
~~3& 0.00&0.39&0.58&0.83&1.00&&0.00&0.75&1.26&1.98&2.43\tabularnewline
~~8& 0.03&0.36&0.58&0.85&1.00&&0.00&0.78&1.27&1.99&2.47\tabularnewline
\hline
{\bfseries Hours}&&&&&&&&&&&\tabularnewline
~~6& 0.00&0.41&0.64&0.90&1.00&&0.00&0.53&1.15&1.97&2.21\tabularnewline
~~12& 0.04&0.41&0.61&0.85&1.00&&0.00&0.87&1.27&1.98&2.33\tabularnewline
~~24&-0.78&0.24&0.45&0.60&1.00&&0.00&0.98&1.40&1.98&2.47\tabularnewline
\hline
{\bfseries Tide}&&&&&&&&&&&\tabularnewline
~~0.25& 0.03&0.35&0.55&0.81&1.00&&0.00&0.81&1.29&1.99&2.47\tabularnewline
~~0.5&-0.10&0.37&0.57&0.82&1.00&&0.00&0.78&1.26&1.97&2.40\tabularnewline
~~1&-0.78&0.36&0.57&0.82&1.00&&0.00&0.76&1.26&1.97&2.40\tabularnewline
\hline
\end{tabular}
\end{center}
\end{table}

@@

% correlations with tide before/after wtreg
% latex.default(tab, file = "", rowlabel = "Site", rgroup = unique(rows),      n.rgroup = rep(2, 4), insert.bottom = foot.val, caption = cap.val,      colheads = c("DO", "Pg\\textsuperscript{\\textit{a}}", "Rt",          "NEM"), caption.loc = "top", rowname = rep(c("Observed",          "Detided"), 4), label = "tab:cor_res") 
%
\begin{table}[!tbp]
\caption{Correlations of tidal changes at each site with continuous \ac{DO} observations, instantaneous \ac{DO} flux, and metabolism estimates (gross production, respiration, and net metabolism) before (observed) and after detiding with weighted regression.  \ac{DO} values are correlated with predicted tidal height at each observation, \ac{DO} flux values are correlated with tidal change between observations, and metabolism estimates are correlated with mean tidal change between observations during day, night, or total, day periods for production, respiration, and net metabolism respectively.  Total day periods were defined as the approximate 24 hours between sunrises for two calendar days, where day was the time from sunrise to sunset and night was the following time from sunset to sunrise.  Results are for weighted regressions using a window of thirty days and tidal height proportion of one.\label{tab:cor_res}} 
\begin{center}
\begin{tabular}{lllll}
\hline\hline
\multicolumn{1}{l}{Site}&\multicolumn{1}{c}{DO}&\multicolumn{1}{c}{Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{Rt}&\multicolumn{1}{c}{NEM}\tabularnewline
\hline
{\bfseries ELKVM}&&&&\tabularnewline
~~Observed& 0.48***& 0.43***& 0.72***& 0.02 \tabularnewline
~~Detided&-0.04***&-0.13*&-0.27***& 0.03 \tabularnewline
\hline
{\bfseries PDBBY}&&&&\tabularnewline
~~Observed&-0.45***&-0.14*&-0.27***&-0.15*\tabularnewline
~~Detided& 0.05***& 0.43***& 0.38***&-0.18**\tabularnewline
\hline
{\bfseries RKBMB}&&&&\tabularnewline
~~Observed& 0.28***& 0.35***& 0.37***& 0.14*\tabularnewline
~~Detided&-0.02**&-0.36***&-0.45***& 0.05 \tabularnewline
\hline
{\bfseries SAPDC}&&&&\tabularnewline
~~Observed& 0.48***& 0.46***& 0.62***& 0.17**\tabularnewline
~~Detided&-0.02*&-0.01 & 0.17**&-0.09 \tabularnewline
\hline
\end{tabular}
\end{center}
\footnotesize *$p<0.05$; **$p<0.01$; ***$p<0.001$\\\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}

% case study metabolism results, including perc anom
% latex.default(tab, file = "", rowlabel = "Site", insert.bottom = foot.val,      caption = cap.val, caption.loc = "top", rgroup = unique(to_tab$Site),      n.rgroup = rep(2, 4), cgroup = c("Pg\\textsuperscript{\\textit{a}}",          "Rt", "NEM"), n.cgroup = c(3, 3, 2), rowname = rows,      colheads = c(rep(c("Mean", "Std. Err.", "Anom"), 2), c("Mean",          "Std. Err.")), label = "tab:case_res") 
%
\begin{table}[!tbp]
\caption{Summary of metabolism estimates (gross production, respiration, and net metabolism) for case studies using \ac{DO} time series before (observed) and after (detided) application of weighted regression.  Means and standard errors are based on daily integrated metabolism estimates.  Anomalous values are the proportion of metabolism estimates that were negative for gross production and positive for respiration.  Results are for weighted regressions with a window of thirty days and tidal height proportion of one.\label{tab:case_res}} 
\begin{center}
\begin{tabular}{llllclllcll}
\hline\hline
\multicolumn{1}{l}{\bfseries Site}&\multicolumn{3}{c}{\bfseries Pg\textsuperscript{\textit{a}}}&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Rt}&\multicolumn{1}{c}{\bfseries }&\multicolumn{2}{c}{\bfseries NEM}\tabularnewline
\cline{2-4} \cline{6-8} \cline{10-11}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}&\multicolumn{1}{c}{Anom}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Std. Err.}\tabularnewline
\hline
{\bfseries ELKVM}&&&&&&&&&&\tabularnewline
~~Observed&5.42&0.70&0.28&&-8.58&0.78&0.23&&-3.16&0.40\tabularnewline
~~Detided&2.81&0.19&0.13&&-2.83&0.21&0.15&&-0.02&0.05\tabularnewline
\hline
{\bfseries PDBBY}&&&&&&&&&&\tabularnewline
~~Observed&8.03&0.70&0.20&&-8.59&0.78&0.16&&-0.56&0.57\tabularnewline
~~Detided&7.82&0.50&0.05&&-7.82&0.50&0.06&& 0.00&0.11\tabularnewline
\hline
{\bfseries RKBMB}&&&&&&&&&&\tabularnewline
~~Observed&2.46&0.14&0.14&&-3.24&0.16&0.10&&-0.78&0.09\tabularnewline
~~Detided&2.61&0.08&0.00&&-3.21&0.08&0.00&&-0.59&0.04\tabularnewline
\hline
{\bfseries SAPDC}&&&&&&&&&&\tabularnewline
~~Observed&4.46&0.25&0.17&&-6.37&0.27&0.11&&-1.90&0.15\tabularnewline
~~Detided&4.37&0.10&0.01&&-5.54&0.11&0.00&&-1.18&0.05\tabularnewline
\hline
\end{tabular}
\end{center}
\textsuperscript{\textit{a}}Pg: gross production, Rt: respiration, NEM: net ecosystem metabolism\end{table}
